Dirac Hamiltonian

### 2017-04-19 Dirac Hamiltonian/dirac_H.tex
[minor] Expressions for Gamma matrices in dim=4n

The Dirac equation aims to change the Klein-Gorden equation into a first order
PDE. There are two approaches to derive the equation. The first can be found in
section 8.2 of \cite{Sakurai2011}, and section II.1 of \cite{zee2010quantum}.
The second approach is found in page 99 (chapter 2) of \cite{Greiner1997}, and
is prefered in my opinion. After all, they lead to the same equation:

@@

In other words, $\gamma^\mu$ is a representation of the Clifford algebra
$\mathrm{Cl}_{1,3}(\R)$. We can cast this equation into another form:
\begin{equation}
    i\pdv{t} ket{\psi} = H \ket{\psi} = \bm{\alpha}\vdot\vb{p}+\beta m
\end{equation}
where $\vb{p}$ is the momentum, $\bm{\alpha}$ and $\beta$ satisfy a similar
relation:
\begin{align}
    \frac{1}{2}\{\alpha_i,\alpha_j\} = \delta_{ij} \\
    \{\alpha_i,\beta\} &= 0 \\
    \alpha_i^2 = \beta^2 &= 0
\end{align}
They are related to $\gamma^\mu$ matrices by:
\begin{equation}
    \alpha_i = \gamma^0 \gamma^i \quad\text{and}\quad
    \beta = \gamma^0
\end{equation}
A detailed formula for $\gamma$ matrices can be found in p.94 of
\cite{zee2010quantum}. A detailed formula for $\alpha_i,\beta$
matrices can be found in p.496 of \cite{Sakurai2011}.

@@

@@

\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

>>>>

The Dirac equation aims to change the Klein-Gorden equation into a first order
PDE. There are two approaches to derive the equation. The first can be found in
section 8.2 of \cite{Sakurai2011}, and section II.1 of \cite{zee2010quantum}.
The second approach is found in page 99 (chapter 2) of \cite{Greiner1997}, and
is preferred in my opinion. After all, they lead to the same equation:

@@

In other words, $\gamma^\mu$ is a representation of the Clifford algebra
$\mathrm{Cl}_{1,3}(\R)$. We can cast this equation into another form:
\begin{equation}
    i\pdv{t} ket{\psi} = H \ket{\psi} = \bm{\alpha}\vdot\vb{p}+\beta m
\end{equation}
where $\vb{p}$ is the momentum, $\bm{\alpha}$ and $\beta$ satisfy a similar
relation:
\begin{align}
    \frac{1}{2}\{\alpha_i,\alpha_j\} = \delta_{ij} \\
    \{\alpha_i,\beta\} &= 0 \\
    \alpha_i^2 = \beta^2 &= 0
\end{align}
They are related to $\gamma^\mu$ matrices by:
\begin{equation}
    \alpha_i = \gamma^0 \gamma^i \quad\text{and}\quad
    \beta = \gamma^0
\end{equation}
A detailed formula for $\gamma$ matrices can be found in p.94 of
\cite{zee2010quantum} or in the section~\todo{}. A detailed formula for $\alpha_i,\beta$
matrices can be found in p.496 of \cite{Sakurai2011}.

@@

\section{Gamma matrices}
\label{sec:Gamma-matrices}

It could be quite easy to find representations of Clifford Algebra in
any dimension. First I note that the size of the Gamma matrices of
$4n\times 4n$, i.e., the dimension could/must
\footnote{I am not sure about this. In this \cite{thePhyStackPage}
    says it must be $4n$, the \cite{theMathematicaPage} says there
could be Gamma matrix in odd dimension} be $4n$. So we could have a
very easy construction by repeatedly using the Pauli matrices. The
Pauli matrices follow:
\begin{equation}
    \{\sigma_i,\sigma_j\} = 2\delta_{ij}
\end{equation}
So the following $4\times 4$ matrices are good Gamma matrices:
\begin{equation}
    \gamma_{1/2/3} = \begin{pmatrix}
                         & \sigma_{1/2/3} \\
        -\sigma_{1/2/3} & \\
    \end{pmatrix},\quad \gamma_0 = \begin{pmatrix}
        \id & \\
           & -\id
    \end{pmatrix}
\end{equation}
Note the minus sign in the $-\id$ of $\gamma_0$ is important to
produce the correct anticommutation relations.

Next, for higher dimension $8$, we can set:
\begin{equation}
    \gamma'_{1/2/3} = \begin{pmatrix}
                         & \gamma_{1/2/3} \\
        -\gamma_{1/2/3} & \\
    \end{pmatrix},\quad \gamma'_0 = \begin{pmatrix}
        \id & \\
           & -\id
    \end{pmatrix}
\end{equation}

And repeating the above process, we can get the Gamma matrices in any
$4n$ dimension.

@@

\begin{thebibliography}{1}
    \bibitem{thePhyStackPage}
    \url{https://physics.stackexchange.com/questions/53318/dimension-of-dirac-gamma-matrices}
    \bibitem{theMathematicaPage}
    \url{http://demonstrations.wolfram.com/DiracMatricesInHigherDimensions/}
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}
